Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-13 paper.qmd
Added Figure labels

mse_two_curvy_b <- ggplot(error_two_curvy_umap, 
                     aes(x = a1, 
                         y = sqrt(MSE))) +
  geom_vline(xintercept = 0.09,
           color = "#d95f02", linetype="dashed", 
           linewidth=1) +
  geom_vline(xintercept = 0.05,
             color = "#1f78b4", 
             linetype="solid",
             linewidth=1) +
  geom_vline(xintercept = 0.03, linetype="dotted",
             color = "#1b9e77", linewidth=1) +
  geom_line(linewidth = 0.5) + 
  geom_point(size = 1) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = "RMSE") +
  #ggtitle("(a)") +
  interior_annotation("a", position = c(0.28, 0.15), cex = 2) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

prop_dens_a1 <- ggplot(error_two_curvy_umap,
                     aes(x = a1,
                         y = prop_comp)) +
  geom_vline(xintercept = 0.09,
           color = "#d95f02", linetype=2, 
           linewidth=1) +
  geom_vline(xintercept = 0.05,
             color = "#1f78b4", 
             linetype="solid",
             linewidth=1) +
  geom_vline(xintercept = 0.03, linetype=2,
             color = "#1b9e77", linewidth=1) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
  ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        axis.title = element_text(size = 18),
axis.text = element_text(size = 15))

@@

a1_m_two_curvy <- ggplot(error_two_curvy_umap,
                     aes(x = a1,
                         y = prop_bins)) +
  geom_vline(xintercept = 0.09,
           color = "#d95f02", linetype=2, 
           linewidth=1) +
  geom_vline(xintercept = 0.05,
             color = "#1f78b4", 
             linetype="solid",
             linewidth=1) +
  geom_vline(xintercept = 0.03, linetype=2,
             color = "#1b9e77", linewidth=1) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression("prop. non-empty bins " * bgroup("(", m / b, ")"))) +
  ggtitle("(b)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        axis.title = element_text(size = 18),
axis.text = element_text(size = 15))
```

@@

mse_two_curvy_lwd <- ggplot(error_rm_two_curvy, 
                     aes(x = benchmark_rm_lwd, 
                         y = MSE,
                         color = bin1)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.5
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  # scale_x_continuous("standardized bin count", 
  #        transform = "reverse") + 
  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 3, 5, 7, 9, 11, 13, 15, 17)]) +
  scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77")) +
  xlab("threshold to remove low-density hexagons") +
  ylab("MSE") +
  ggtitle("(c)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.title = element_text(size = 18),
axis.text = element_text(size = 15))
```

>>>>

mse_two_curvy_b <- ggplot(error_two_curvy_umap, 
                     aes(x = a1, 
                         y = sqrt(MSE))) +
  geom_vline(xintercept = 0.09,
           color = "#d95f02", linetype="dashed", 
           linewidth=1) +
  geom_vline(xintercept = 0.05,
             color = "#1f78b4", 
             linetype="solid",
             linewidth=1) +
  geom_vline(xintercept = 0.03, linetype="dotted",
             color = "#1b9e77", linewidth=1) +
  geom_line(linewidth = 0.5) + 
  geom_point(size = 1) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = "RMSE") +
  interior_annotation("a", position = c(0.9, 0.15), cex = 2) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

prop_dens_a1 <- ggplot(error_two_curvy_umap,
                     aes(x = a1,
                         y = prop_comp)) +
  geom_vline(xintercept = 0.09,
           color = "#d95f02", linetype=2, 
           linewidth=1) +
  geom_vline(xintercept = 0.05,
             color = "#1f78b4", 
             linetype="solid",
             linewidth=1) +
  geom_vline(xintercept = 0.03, linetype=2,
             color = "#1b9e77", linewidth=1) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = paste("Relative proportion density")) +
  interior_annotation("d", position = c(0.9, 0.9), cex = 2) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        axis.title = element_text(size = 18),
axis.text = element_text(size = 15))

@@

a1_m_two_curvy <- ggplot(error_two_curvy_umap,
                     aes(x = a1,
                         y = prop_bins)) +
  geom_vline(xintercept = 0.09,
           color = "#d95f02", linetype=2, 
           linewidth=1) +
  geom_vline(xintercept = 0.05,
             color = "#1f78b4", 
             linetype="solid",
             linewidth=1) +
  geom_vline(xintercept = 0.03, linetype=2,
             color = "#1b9e77", linewidth=1) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1) +
  # scale_y_continuous(breaks = sort(unique(error_two_curvy$b_non_empty))[-3]) +
  scale_x_continuous(breaks = sort(unique(round(error_two_curvy_umap$a1, 2)))[c(1, 4, 7, 11, 13, 14, 15, 16)]) +
  labs(x = expression(paste("binwidth (", a[1], ")")), y = expression("prop. non-empty bins " * bgroup("(", m / b, ")"))) +
  interior_annotation("b", position = c(0.9, 0.15), cex = 2) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        axis.title = element_text(size = 18),
axis.text = element_text(size = 15))
```

@@

mse_two_curvy_lwd <- ggplot(error_rm_two_curvy, 
                     aes(x = benchmark_rm_lwd, 
                         y = sqrt(MSE),
                         color = bin1)) + 
  geom_point(
    size = 1
    ) +
  geom_line(
    linewidth = 0.5
    ) + 
  # geom_vline(xintercept = benchmark1, linetype="solid", 
  #            color = "#bdbdbd", linewidth=1, alpha = 0.5) +
  # scale_x_continuous("standardized bin count", 
  #        transform = "reverse") + 
  scale_x_continuous(breaks = sort(unique(error_rm_two_curvy$benchmark_rm_lwd))[c(1, 3, 5, 7, 9, 11, 13, 15, 17)]) +
  scale_color_manual(values=c("#d95f02", "#1f78b4", "#1b9e77")) +
  xlab("threshold to remove low-density hexagons") +
  ylab("RMSE") +
  interior_annotation("c", position = c(0.1, 0.9), cex = 2) +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 20, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.title = element_text(size = 18),
axis.text = element_text(size = 15))
```
